Specifgic Gene or Metabolite Levels

### 2025-01-13 OmicsQuickPlotFunctions.Rmd
updated raw metabolomic data

library(plotly)
library(edgeR)

GeneIDKey = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/GeneralDataFiles/GeneIDKey.csv", row.names = 1)

##load all files from github
git.dir = "https://raw.githubusercontent.com/johncsantiago/WhartonLab/master/"

TKT.cpm = read.csv(paste0(git.dir,"TKT_RNAseq/CountTables/TKT_cpmdata.csv"),row.names = 1)

TKT.groups  = read.csv(paste0(git.dir,"TKT_RNAseq/CountTables/TKT.metadata.csv"),row.names = 1)
TKT.FDR = read.csv(paste0(git.dir,"TKT_RNAseq/CountTables/TKT.EdgeR.FDRTable.csv"),row.names = 1)
TKT.FC = read.csv(paste0(git.dir,"TKT_RNAseq/CountTables/TKT.EdgeR.FCTable.csv"),row.names = 1)

@@

A4V.cpm = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/A4V.cpmdata.csv",row.names = 1)

A4V.FDR = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/A4V.FDRdata.csv")

gbbOE = read.csv("/Users/johncsantiago/Documents/GitHub/Datasets/gbbOE_CountTable.csv", row.names = 1)
gbbOE.groups = read.csv("/Users/johncsantiago/Documents/GitHub/Datasets/gbbOE_Metadata.csv", row.names = 1)
gbbOE = gbbOE[,row.names(gbbOE.groups)]
gbbOE.groups$group = paste0(gbbOE.groups$genotype, "_",gbbOE.groups$stage)

##normalize data
countdata=gbbOE
x <- countdata
gr <- factor(gbbOE.groups$group, levels=c(unique(gbbOE.groups$group)))
y <- DGEList(counts=x,group=gr)
keep <- filterByExpr(y)
y <- y[keep,,keep.lib.sizes=FALSE]
z <- calcNormFactors(y, method = "TMM")
##normalized cpm
gbbOE.cpmdata=cpm(z)

gbbOEmean = gbbOE.cpmdata[,1:6]
colnames(gbbOEmean) = unique(gr)
i=1
while(i<=6){
  temp= gbbOE.cpmdata[,row.names(gbbOE.groups[gbbOE.groups$group == colnames(gbbOEmean)[i],])]
  gbbOEmean[,i] = apply(temp, 1, mean)
  i=i+1
}

design<-model.matrix(~0 + gr)
colnames(design) = unique(gr)
z = estimateDisp(z, design, robust = T)

fit <- glmQLFit(z, design)
plotQLDisp(fit)

compare = makeContrasts(G85R_LL3-LoxP_LL3, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GRxWT.LL3=DEout
sGRxWT.LL3=DEout[DEout$FDR<.05,]

compare = makeContrasts(G85R_ML3-LoxP_ML3, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GRxWT.ML3=DEout
sGRxWT.ML3=DEout[DEout$FDR<.05,]

compare = makeContrasts(G85R_LL3-G85R_ML3, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GR.LL3xML3=DEout
sGR.LL3xML3=DEout[DEout$FDR<.05,]

compare = makeContrasts(LoxP_LL3-LoxP_ML3, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
WT.LL3xML3=DEout
sWT.LL3xML3=DEout[DEout$FDR<.05,]

compare = makeContrasts(G85R_gbb_LL3-G85R_gbb_ML3, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
G85R_gbb.LL3xML3=DEout
sG85R_gbb.LL3xML3=DEout[DEout$FDR<.05,]

gbbKO = read.csv("/Users/johncsantiago/Documents/GitHub/Datasets/gbbKO_CountTable.csv", row.names = 1)
gbbKO.groups = read.csv("/Users/johncsantiago/Documents/GitHub/Datasets/gbbKO_Metadata.csv", row.names = 1)

##normalize data
countdata=gbbKO
x <- countdata
gr <- factor(gbbKO.groups$Genotype, levels=c(unique(gbbKO.groups$Genotype)))
y <- DGEList(counts=x,group=gr)
keep <- filterByExpr(y)
y <- y[keep,,keep.lib.sizes=FALSE]
z <- calcNormFactors(y, method = "TMM")
##normalized cpm
gbbKO.cpmdata=cpm(z)

rawdata = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/Metabolomics/MetabolomicsForMetaboanalyst.txt", sep = "\t", row.names = 1)
KEGG.Key = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/Metabolomics/RawMetabolomicData.csv", row.names = 1)
KEGG.Key = setNames(KEGG.Key[-c(1:2), 2], row.names(KEGG.Key)[-c(1:2)])

@@

@@

```{r echo = F}
##Function to plot all G85R groups

plot.G85R = function(ID){
ID.row = GeneIDKey[grep(ID, GeneIDKey$FBgn),]

@@

```{r echo = F}
##Function to plot a specific gene expression levels in all A4V groups

plot.A4V = function(ID){
ID.row = GeneIDKey[grep(ID, GeneIDKey$FBgn),]

if(nrow(ID.row) == 1){
  FBgn = ID.row$FBgn
  name = ID.row$Symbol
  A4V.groups = as.vector(colnames(A4V.cpm))

@@

```{r echo = F}
##Plot levels of a specific metabolite

plot.metab = function(metabID){

ID.row = row.names(norm.data)[grep(ID, row.names(norm.data))]

if(length(ID.row) == 1){
metab = ID.row
name = ID.row

@@

@@

@@

@@

ID = "0000018"
plot.G85R(ID)

@@

```{r echo = F}
##Tailared code to plot a specific gene expression levels in all A4V groups

ID = "0000018"

@@

metabID = 'NADP'

@@

```{r echo = F}
##Tailared code to plot a subset of groups in the TKT dataset

@@

@@

if(nrow(ID.row) != 1){
  ID.row
}

@@

@@

if(nrow(ID.row) != 1){
  ID.row
}
```

>>>>

##library(plotly)
library(edgeR)

##load all files from github
git.dir = "https://raw.githubusercontent.com/johncsantiago/WhartonLab/refs/heads/master/GeneralDataFiles/"

GeneIDKey = read.csv(paste0(git.dir, "GeneIDKey.csv"), row.names = 1)

TKT.cpm = read.csv(paste0(git.dir,"TKT_cpmdata.csv"), row.names = 1)

TKT.groups  = read.csv(paste0(git.dir, "TKT.metadata.csv"), row.names = 1)

@@

A4V.cpm = read.csv(paste0(git.dir, "A4V.cpmdata.csv"), row.names = 1)

gbbOE = read.csv(paste0(git.dir, "gbbOE_CountTable.csv"), row.names = 1)

gbbOE.groups = read.csv(paste0(git.dir, "gbbOE_Metadata.csv"), row.names = 1)

gbbOE = gbbOE[,row.names(gbbOE.groups)]
gbbOE.groups$group = paste0(gbbOE.groups$genotype, "_",gbbOE.groups$stage)

rawdata = read.csv(paste0(git.dir, "MetabolomicsForMetaboanalyst.txt"), sep = "\t", row.names = 1)

KEGG.Key = read.csv(paste0(git.dir, "RawMetabolomicData.csv"), row.names = 1)

KEGG.Key = setNames(KEGG.Key[-c(1:2), 2], row.names(KEGG.Key)[-c(1:2)])

@@

SCdata = read.csv(paste0(git.dir, "Nguyen%20Serpe%202024%20scRNA-seq%20VNC.csv"))

@@

```{r echo = F}
##Function to plot all G85R groups
plot.G85R = function(ID){
ID.row = GeneIDKey[c(grep(ID, GeneIDKey$FBgn),
                     grep(ID, GeneIDKey$Symbol)),]

if(nrow(ID.row) < 50 & nrow(ID.row) > 0){
if(nrow(ID.row) == 1){
FBgn = ID.row$FBgn
name = ID.row$Symbol

@@

if(nrow(ID.row) != 1){
  ID.row
}
}
}

##Function to plot a specific gene expression levels in all A4V groups
plot.A4V = function(ID){
ID.row = GeneIDKey[c(grep(ID, GeneIDKey$FBgn),
                     grep(ID, GeneIDKey$Symbol)),]

if(nrow(ID.row) < 50 & nrow(ID.row) > 0){
if(nrow(ID.row) == 1){
  FBgn = ID.row$FBgn
  name = ID.row$Symbol
  A4V.groups = as.vector(colnames(A4V.cpm))

@@

if(nrow(ID.row) != 1){
  ID.row
}
}
}

##Plot levels of a specific metabolite
plot.metab = function(metabID){

ID.row = grep(metabID, row.names(norm.data))

if(length(ID.row) < 50 & length(ID.row) > 0){
if(length(ID.row) == 1){
metab = row.names(norm.data)[ID.row]
name = row.names(norm.data)[ID.row]

@@

rect(0, 0-(1.5*max(na.omit(metab$data))), 13, 1.5*max(na.omit(metab$data)), col = 'gray95')

@@

}
}

##Function to plot a specific gene expression levels in all cell types for the SCdata
plot.SC = function(gene.name){
ID.row = SCdata[grep(gene.name, SCdata$Gene.name),]

if(nrow(ID.row) < 50 & nrow(ID.row) > 0){
if(nrow(ID.row) == 1){

  gene.name = ID.row$Gene.name
  gene.levels = setNames(as.numeric(ID.row[,colnames(ID.row) != "Gene.name"]), colnames(ID.row)[colnames(ID.row) != "Gene.name"]) 
  gene.levels = gene.levels[18:1]
  highlight = rep(0, 18)
  highlight[(gene.levels > 2*(mean(gene.levels)))] = 1
  highlight = gene.levels * highlight

  nochange = gene.levels
  nochange[gene.levels > 1.5] = 0
  nochange = gene.levels * nochange

  label.colors = highlight

  label.colors[label.colors > 0] = "deeppink"
  label.colors[label.colors == 0] = "black"
  label.colors[nochange > 0] = 'steelblue'

  if(1.25*(max(gene.levels)) < 5){
    xmax = 5
  }

  if(1.25*(max(gene.levels)) >= 5){
    xmax = 1.25*(max(gene.levels))
  }

  par(mar=c(5, 12, 4, 6), xpd=TRUE)

  plot(x = NA,
      y = NA,
     type = 'n',
     ylim = c(.95, 21),
     xlim = c(0, xmax),
     xaxt = "n",
     ylab="",
     yaxt = "n",
     xlab = "Mean UMI Counts",
     bty = "n",
     main = gene.name)

  barplot(height = gene.levels, 
          horiz = T, 
          #las = 2,
          xlim = c(0, xmax),
          names.arg = NA,
          border = T,
          add = T)

    barplot(height = highlight, 
          horiz = T, 
          #las = 2, 
          names.arg = NA,
          border = T,
          col = "deeppink",
          add = T)

    barplot(height = nochange, 
          horiz = T, 
          #las = 2, 
          names.arg = NA,
          border = T,
          col = "lightsteelblue",
          add = T)

  mtext(side = 2,
        las = 2,
        at = .7+(1.2*(0:17)),
        text = names(gene.levels),
        col = label.colors,
        line = 0)

  abline(v = 1,
         lty = 2,
         lwd = 2,
         xpd = F)

  legend(x = .9*xmax,
         y = 23,
         legend = c("high", "mid", "low"),
         fill = c("deeppink", "grey", "lightsteelblue"),
         bty = "n",
         xpd = T)

@@

if(nrow(ID.row) != 1){
  ID.row
}
}
}

@@

## Plot gene expression from G85R TKT RNAseq experiment
## ID needs to be the full or partial gene symbol or FBgn (case sensitive)

ID = "Arc1"

plot.G85R(ID)

@@

## Plot gene expression from A4V RNAseq experiment
## ID needs to be the full or partial gene symbol or FBgn (case sensitive)

ID = "Arc1"

@@

## Plot metabolite levels from the G85R TKT metabolomics experiment
## metabID needs to be the full or partial metabolite name as listed in the table (case sensitive)

metabID = 'NADH'

@@

## Plot gene expression from Single Cell RNAseq experiment
## gene name needs to be the full or partial gene symbol as listed in the table (case sensitive)

gene.name = 'gbb'

plot.SC(gene.name)

```{r echo = F}
## Plot a subset of groups in the TKT dataset

@@

if(nrow(ID.row) < 500 & nrow(ID.row) > 0){
if(nrow(ID.row) == 1){
FBgn = ID.row$FBgn
name = ID.row$Symbol

@@

@@

if(nrow(ID.row) < 500 & nrow(ID.row) > 0){

@@
